Cores from kicks: constraints for observations of gravitational recoil in elliptical galaxies

### 2023-09-22 papers/paper-corekick/paper-corekick.tex
Add text about simulation details

\newcommand{\Msun}{\ensuremath{M_{\sun}}}

@@

\drafting{Rantala 2018 ICs}
\begin{equation}\label{eq:dehnen}
    \rho_\star(r) = \frac{(3-\gamma)M_\star}{4\pi} \frac{a}{r^\gamma (r+a)^{(4-\gamma)}}.
\end{equation}

\drafting{We then construct isolated merger initial conditions}. 
Interactions between stellar particles are softened with a softening length of \drafting{$\varepsilon = 2.5\,\mathrm{pc}$, and the \ketju{} region radius is set to $r_{\rm ketju}=3\varepsilon = 7.5\,\mathrm{pc}$.}

% RESULTS
\section{Results}\label{sec:results}
\drafting{
    Things to show:
    \begin{enumerate}
        \item DAG for hierarchical model
        \item Density profiles 
        \item Scatter plot $v_k\mathrm{--}r_b$ with best fit function
        \item Relation $v_k\mathrm{--}r_b$ model comparison (elpd?)
        \item MC-sampled distribution of $r_b$
        \item $h_4$ maps
    \end{enumerate}
}

>>>>

\newcommand{\Msun}{\ensuremath{\mathrm{M}_{\sun}}}
\newcommand{\kmps}{\ensuremath{\mathrm{km}\,\mathrm{s}^{-1}}}

@@

\drafting{Rantala 2018 ICs}
We model the merger of two gas-poor elliptical galaxies using collisionless merger simulations.
Our galaxy initial conditions are chosen to match the model IC-3 presented in \citet{rantala2019}. 
The galaxy is thus represented as a multicomponent sphere, including a stellar component of total mass $M_\star\sim 1.38\times10^{11}\,\Msun$ embedded within a DM \drafting{ensure abbreviation stated} component of total mass $M_\mathrm{DM}=2.5\times10^{13}\,\Msun$, and at the centre a SMBH with mass $M_\bullet=2.93\times10^{9}\,\Msun$.
The stellar and DM components each follow a \citet{dehnen1993} profile with shape parameter $\gamma=1$, and scale radius $a_\star=3.9\,\mathrm{kpc}$ and $a_\mathrm{DM}=245\,\mathrm{kpc}$, respectively\footnote{A Dehnen profile with $\gamma=1$ is commonly referred to as a Hernquist profile.}.
The density profile is given by:
\begin{equation}\label{eq:dehnen}
    \rho_\star(r) = \frac{(3-\gamma)M_\star}{4\pi} \frac{a}{r^\gamma (r+a)^{(4-\gamma)}}.
\end{equation}
The mass of a stellar particle is set to $m_\star=5\times10^4\,\Msun$, and the mass of a DM particle is set to $m_\mathrm{DM}=5\times10^6\,\Msun$.

We create six independently Monte Carlo sampled galaxy ICs, and merge each combination of galaxies on a radial orbit.
The merger orbit has an initial separation of $D=30\,\mathrm{kpc}$, a first pericentre distance of $r_\mathrm{peri}=2\,\mathrm{kpc}$, and an initial eccentricity of $e_0=0.97$.
We select a merger combination that results in a rapid coalescence of the BH binary, noting that the merger timescale we observe is driven by stochasticity in the binary eccentricity \citep{nasim2020,rawlings2023}.
We then select the snapshot just prior to merger, and generate 31 `child' simulations, where each child has a unique gravitational recoil kick velocity $v_\mathrm{kick}$ prescribed, ranging from $0\,\kmps$ to $1800\,\kmps$ (inclusive), in $60\,\kmps$ increments.
The upper limit on $v_\mathrm{kick}$ is chosen to match the escape velocity of the centre of the merger remnant.
We additionally run one simulation with a recoil kick above the escape velocity, with $v_\mathrm{kick}=2000\,\kmps$.
Each child simulation is run until the median remnant SMBH velocity relative to the galaxy CoM \drafting{check abbreviation} velocity is less than $10\,\kmps$ for at least $0.1\,\mathrm{Gyr}$.
Interactions between stellar particles are softened with a softening length of $\varepsilon = 2.5\,\mathrm{pc}$, and the \ketju{} region radius is set to $r_{\rm ketju}=3\varepsilon = 7.5\,\mathrm{pc}$.

\drafting{Atte: mergers of unequal mass ratio?}

% RESULTS
\section{Results}\label{sec:results}
\drafting{
    Things to show:
    \begin{enumerate}
        \item DAG for hierarchical model
        \item Density profiles 
        \item Scatter plot $v_k\mathrm{-}r_b$ with best fit function
        \item Relation $v_k\mathrm{-}r_b$ model comparison (elpd?)
        \item MC-sampled distribution of $r_b$
        \item $h_4$ maps
    \end{enumerate}
}
